Type Error Telemetry at Scale

### 2023-09-30 programming23/paper.tex
more sec 6 cleanup and misc

% TODO BEFORE SUBMISSION
% be consistent between "IDs" and "ids" [SK prefers "IDs" to avoid it being read as a word]

@@

Our approach to data analysis is inspired by Marceau
et~al.~\cite{mfk-onward-2011,mfk-sigcse-2011}, who used keystroke-level data
collected over 6 weeks to analyze students' responses to runtime errors
in DrRacket~\cite{fcffksf-jfp-2002}.
Similarly, Macedo et~al.~\cite{mcpcsprs-abz-2020,mcpcsprs-scp-2021} capture
edit sessions from the online Alloy4Fun editor and present several automated
analyses.
A major challenge in our setting, however, is that edit sequences are
incomplete---there may be minutes-long gaps between two telemetry records.
Several more-recent works have studied edit sequences
to draw conclusions about errors and intent, e.g.,~\cite{wk-koli-2020,lgk-pj-2022,rsgl-cpp-2020}.

The BitFit project collects data about beginning students interactions with a
Java IDE~\cite{ekc-wccce-2016,anna-russo-kennedy-ms-2006}; however, unlike Blackbox, it does
not collect source code.
BitFit data consists of a timestamp, a session id, and counter values showing
the number of attempts at a question, the number of hints requested, the number
of compiles and runs, and the number of submission attempts.
On one hand, this data is more informative than Roblox Studio's because it
targets specific events rather than random keystrokes.
On the other hand, there are no details about errors to show what students
struggled with.
Other logging-based studies of Java programs use the full error message to
draw conclusions~\cite{bgimgm-cse-2016,dlc-iticse-2014}.
Roblox Studio cannot follow this approach because strings in errors may leak
personal information.

Much work on the benefits of static types focuses on a small number
of case studies~\cite{w-popl-1986,hw-scp-2004,td-tosem-2001},
inverviews~\cite{cdhhjklwya-hatra-2020,gstf-hatra-2021,cams-oopsla-2020},
or static corpuses~\cite{rppd-fse-2014,bhmvv-toplas-2019,bmvv-arxiv-2019}
(for literature reviews, see~\cite{empirical-types,heeren-thesis}).
Controlled experiments, for example comparing statically-typed Java and dynamically-typed
Groovy~\cite{khrts-icpc-2012}, are a notable exception.
Lerner et~al.~\cite{lfgc-pldi-2007} apply type error repair tool on over two thousand
of programs written during five homework exercises.
Seidel et~al.~\cite{sscwj-oopsla-2017,sjw-jfp-2018} use five thousand programs
to train data-driven methods for localizing errors.
These large studies with several thousands of programs are
nevertheless on a much smaller scale than Roblox Studio.

@@

%% -----------------------------------------------------------------------------
\acks

Thanks to Benjamin Chung for several helpful discussions about data analysis
and effective plotting.
Greenman was supported by
NSF grant \href{https://nsf.gov/awardsearch/showAward?AWD_ID=2030859&HistoricalAwards=false}{2030859}
to the CRA for the \href{https://cifellows2020.org}{CIFellows} project.

>>>>

% TODO BEFORE SUBMISSION
% be consistent between "IDs" and "ids" [SK prefers "IDs" to avoid it being read as a word]
% even more so between "ID" and "id" [NB: "id" is a word in English]
% \usepackage[sort]{natbib} or otherwise manually fix cites to be in ascending order

@@

Our approach to data analysis is inspired by Marceau
et~al.~\cite{mfk-onward-2011,mfk-sigcse-2011} and
Macedo et~al.~\cite{mcpcsprs-abz-2020,mcpcsprs-scp-2021}
for Racket and Alloy4Fun, respectively. Other more recent work has
also studied edit sequences to infer intent~\cite{wk-koli-2020,lgk-pj-2022,rsgl-cpp-2020}.
A major difference is that we have only samples, not full sequences,
which significantly impacts our study methods.\todo{Reasonable rewrite?}

The BitFit project collects data about beginning students interactions in
Java~\cite{ekc-wccce-2016,anna-russo-kennedy-ms-2006}; it does
not collect source code.\todo{Does it sample? If not, say so!}
However, its data do not contain details about errors.\todo{Really?!?}
Other logging-based studies of Java programs use the full error message to
draw conclusions~\cite{bgimgm-cse-2016,dlc-iticse-2014}, which we
cannot use to preserve privacy.

Much work~\cite{empirical-types,heeren-thesis} on the benefits of static types focuses on
small case studies~\cite{w-popl-1986,hw-scp-2004,td-tosem-2001},
inverviews~\cite{cdhhjklwya-hatra-2020,gstf-hatra-2021,cams-oopsla-2020},
or static
corpuses~\cite{rppd-fse-2014,bhmvv-toplas-2019,bmvv-arxiv-2019}.\todo{Why
  are we citing arXiv?}
Controlled experiments are rare;
Kleinschmager, et~al.\cite{khrts-icpc-2012} is a notable exception.
Lerner et~al.~\cite{lfgc-pldi-2007} apply a type error repair tool on
over two thousand programs.
Seidel et~al.~\cite{sscwj-oopsla-2017,sjw-jfp-2018} use five thousand programs
to train data-driven methods for localizing errors.
These are all on a vastly smaller scale than our study and also study
different issues.

@@

%% -----------------------------------------------------------------------------
% \acks

% Thanks to Benjamin Chung for several helpful discussions about data analysis
% and effective plotting.
% Greenman was supported by
% NSF grant \href{https://nsf.gov/awardsearch/showAward?AWD_ID=2030859&HistoricalAwards=false}{2030859}
% to the CRA for the \href{https://cifellows2020.org}{CIFellows} project.
